---
title: "homework2_sk4833"
author: "Kexin Sun"
date: "9/29/2020"
output: html_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1

First, define a path to the dataset. 

```{r}
path_to_data = "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx"
```


Read the Mr. Trashwheel dataset. 

```{r}
trashwheel_df = 
	read_xlsx(
		path = path_to_data,
		sheet = "Mr. Trash Wheel",
		range = cell_cols("A:N")) %>% 
	janitor::clean_names() %>% 
	drop_na(dumpster) %>% 
	mutate(
		sports_balls = round(sports_balls),
		sports_balls = as.integer(sports_balls)
	)
```

Read precipitation data! For 2018 and 2017. 

```{r}
precip_2018 = 
	read_excel(
		"./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "2018 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2018) %>% 
	relocate(year)
precip_2017 = 
	read_excel(
		"./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "2017 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2017) %>% 
	relocate(year)
```

Now combine annual precipitation dataframes. In the following code chunk, I create a "helper" tibble that contains pairs of numeric and character ways of representing month, and then merge that (using month number as a key) with the precipitation dataset. This technique is one I use often when I need to recode a moderate or large number of values for a variable. 

```{r}
month_df = 
	tibble(
		month = 1:12,
		month_name = month.name
	)
precip_df = 
	bind_rows(precip_2018, precip_2017)
precip_df =
	left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data. In this dataset:

* The median number of sports balls found in a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`
* The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.


## Problem 2



Read the data.


```{r}
subway_df = 
  read_csv ("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  select(-division, -exit_only, -staffing, -staff_hours, -ada_notes, -free_crossover, 
         -north_south_street, -east_west_street, -corner, -entrance_latitude,   
         -entrance_longitude,
         -station_location, -entrance_location) %>%
  mutate(entry = recode (entry, 'YES' = TRUE , 'NO' = FALSE ))
```


After I read the data from **data** folder in my computer, I use R function janitor::clean_names() to clean the name of each variable, and also, some other unimportant variables, such as **division** and **staffing**, are omitted in this project

There are `r ncol(subway_df)` variables and `r nrow(subway_df)` observations in this data set, and variables are "`r names(subway_df)`".

However, this data is still not tidy yet.


There are 465 distinct stations and 84 ADA compliant stations.

```{r}
subway_df2= 
  distinct(subway_df, station_name, line, ada) %>% 
  filter(ada == TRUE) %>% 
  nrow()
```


There are 183 station entrances/exits without vending, and there are 69 of them allow entrance, so the proportion is 69/183 = 0.377.


```{r}

subway_df3= 
  select(subway_df, entry, vending) %>% 
  filter(vending == "NO") %>% 
  nrow()

subway_df4= 
  select(subway_df, entry, vending) %>% 
  filter(vending == "NO") %>% 
  filter(entry == TRUE) %>% 
  nrow()
```



Since variables Route8 to Route9 are double. We need to firstly convert them to character.

And then use pivot_longer function to generate two new variables called **route_name** and **route_number**.


```{r}
subway_df_clean = 
  mutate_at(subway_df, vars(route8:route11), as.character) %>%
  pivot_longer(
    route1:route11,
    names_to = "route_name",
    values_to = "route_number"
  ) 
```


There are 60 distnict stations served A train, and there are 17 of them are ADA compliant. 

```{r}
filter(subway_df_clean, route_number == "A") %>% 
  select(route_number, station_name, line) %>% 
  distinct(route_number, station_name, line)


filter(subway_df_clean, route_number == "A", ada == "TRUE") %>% 
  select(route_number, station_name, ada, line) %>%  
  distinct(route_number,station_name, ada, line) 
```




## Problem 3

Read the pols_month data and clean it.

```{r}
pols_month_df = 
          read_csv("./data/pols-month.csv") %>% 
          janitor::clean_names() %>% 
          separate(mon, c("year", "month", "day"), sep = "([-])") %>% 
          mutate(month = as.numeric(month), 
                 day = as.numeric(day),      ## convert day,year,month to numerical variable
                 year = as.numeric(year)) %>% 
                 
          relocate(prez_gop, prez_dem) %>% 
  
          pivot_longer(                            ## take values of gop and dem
                 prez_gop:prez_dem,
                 names_to = "president",           ## create a variable called president
                 names_prefix = "prez_",
                 values_to = "number") %>% 
  
          filter(number == 1) %>% 
          select(-day) 
          
          
month_df = 
  tibble(
    month = 1:12, 
    month_name = month.name
    )

pols_month_df = 
         left_join(pols_month_df, month_df, by = "month") %>% 
         relocate(year, month_name) %>% 
         select(-month) %>% 
         rename(month = month_name)
```




Read the snp data set and clean it.

```{r}
snp_df = 
  read_csv ("./data/snp.csv") %>%
  janitor::clean_names() %>% 
  separate(date, c("month", "day", "year"), sep = "([/])") %>% 
            mutate(month = as.numeric(month),
                 day = as.numeric(day),
                 year = as.numeric(year)
                ) %>% 
          select(-day)
        
          
month_df = 
  tibble(
    month = 1:12, 
    month_name = month.name
    )

snp_df = 
         left_join(snp_df, month_df, by = "month") %>% 
         relocate(year, month_name) %>% 
         select(-month) %>% 
         rename(month = month_name)
```


Read the unemployment data set and clean it.

```{r}
unemployment_df = 
  read_csv ("./data/unemployment.csv") %>%
  janitor::clean_names() %>% 
  pivot_longer(
              jan:dec,
              names_to = "month",
              values_to = "unemployment rate"
              ) %>% 
  mutate(month = recode(month, "jan" = "January", "feb" = "February", "mar" = "March", 
                        "apr" = "April", "may" = "May", "jun" = "June", "jul" = "July",
                        "aug" = "August", "sep" = "September", "oct" = "October",
                        "nov" = "November", "dec" = "December"))    ## Rename month name


```

Finally, we merge these three data set together by using variables month and year.

```{r}
merge_df = 
        left_join(pols_month_df, snp_df, by = c("month", "year")) %>% 
        left_join(., unemployment_df,by = c("month", "year"))
        
```


Short paragraph about these data set.

For pols-month data set, there are `r ncol(pols_month_df)` variables and `r nrow(pols_month_df)` observations in this data set, and variables are "`r names(pols_month_df)`". 

For snp data set, there are `r ncol(snp_df)` variables and `r nrow(snp_df)` observations in this data set, and variables are "`r names(snp_df)`". 

For unemployment data set, there are `r ncol(unemployment_df)` variables and `r nrow(unemployment_df)` observations in this data set, and variables are "`r names(unemployment_df)`". 

For final merged data set, there are `r ncol(merge_df)` variables and `r nrow(merge_df)` observations in this data set, and variables are "`r names(merge_df)`". The range of year is (`r range(pull(merge_df, year))`).
